---
title: "Revisiting the visualizations from The Economist"

output:
  pdf_document: default
  html_document: default
  word_document: default
---

Source :https://medium.economist.com/mistakes-weve-drawn-a-few-8cdd8a42d368"


### *1. FACEBOOK*

```{r pressure, echo=FALSE, fig.height=3, fig.width=6, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(forcats)
facebook <- read_csv('Economist_corbyn.csv')
names(facebook) <- c("page","avglike")
facebook %>%
  ggplot(aes(x=fct_reorder(page,avglike),y=avglike))+
  geom_bar(stat = 'identity', fill = 'red') + 
  labs(x= NULL, y= NULL, title= 'Left-Click',
  subtitle= "Average number of likes per Facebook post\n2016, '000",
  caption = 'Source: Facebook')+
  scale_y_continuous(labels = unit_format(scale = 1e-3, unit = ''),
                     breaks = seq(0, 6000, 1000),
                     limits = c(0, 6000)) +
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.05),
        axis.text.x = element_text(hjust = 0.5))+
  coord_flip()
```

### *2. DOGS MEASUREMENTS*
## **Fit as a Butcher's Dog**
*Characterstics of dogs registered with the UK's Kennel Club,Average when fully grown*

```{r pressure2, echo=FALSE, fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
dogs <- read_csv('Economist_dogs.csv')
names(dogs) <- c("year","wght","neck")
Plot1 <-dogs %>%
  ggplot(aes(x=year,y=wght))+
  geom_line(colour = 'red',size=1.3 ) + 
  labs(x= NULL, y= NULL, title= "Weight* (in kg)",
       caption = '*Where atleast 50 are registered per year')+
  scale_x_continuous(breaks = seq(2006, 2015, 3),
                     limits = c(2006, 2015)) +
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.05),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=14, face="bold.italic"),
        axis.text.x = element_text(hjust = 0.5))
Plot1
Plot2 <-dogs %>%
  ggplot(aes(x=year,y=neck))+
  geom_line(colour = 'red',size=1.3 ) + 
  labs(x= NULL, y= NULL, title= "Neck Size** (in cm)",
       caption = '**Where atleast 100 are registered per year')+
  scale_x_continuous(breaks = seq(2006, 2015, 3),
                     limits = c(2006, 2015)) +
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.05),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=14, face="bold.italic"),
        axis.text.x = element_text(hjust = 0.5))
Plot2
```


Sources: Kennel Club;The Economist    

### *3. OPINION ON BREXIT *
```{r pressure3, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
brexit <- read_csv('Economist_brexit.csv')
names(brexit) <- c("date","right","wrong")
brexit1 <- brexit %>%
  gather(vote,percent,right:wrong)
plot3 <-brexit1 %>%
  ggplot(aes(x=date,y=percent,colour=vote))+
  geom_point(stat='identity', position='identity') +
  geom_smooth(se=FALSE) +
  scale_colour_brewer(type = 'qual', palette = 6, direction =-1)+
  labs(x= NULL, y= "% responding", title= 'Bremorse',
       subtitle= "In hindsight, Do you think Britain was right or wrong to leave the EU?",
       caption = 'Source:NatCen Social Research')+
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.05),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=12, face="bold"),
        axis.text.x = element_text(hjust = 0.5), 
        legend.title = element_blank(),
        legend.position = "top",
        legend.background = element_rect(fill = '#ffef99'),
        legend.key.size=unit(0.04, "cm"))
plot3
```

### 4. *TRADE DEFICIT AND MANUFACTURING EMPLOYMENT *
## **Free Markets and Free Workers**
*United States*

```{r pressure4, echo=FALSE, fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
trade <- read_csv('Economist_us-trade-manufacturing.csv')
names(trade) <- c("year","deficit","empl")
trade <- trade %>%
drop_na()
trade$year <- as.numeric(trade$year)
plot4a <-trade %>%
  ggplot(aes(x=year,y=deficit))+
  geom_bar(stat='identity', position='identity',fill='red') +
  geom_smooth(se=FALSE) +
  labs(x= NULL, y=NULL, title= "Trade Deficit with China in goods \n$bn")+
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.05),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=12, face="bold"),
        axis.text.x =element_text(angle=45))
plot4a
plot4b <-trade %>%
  ggplot(aes(x=year,y=empl))+
  geom_bar(stat='identity', position='identity', fill='red') +
  geom_smooth(se=FALSE) +
  labs(x= NULL, y=NULL, title= "Manufacturing Employment \nm")+
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.05),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=12, face="bold"),
        axis.text.x =element_text(angle=45))
plot4b
```


Sources: US Census Bureau, BLS 

### 5. *POPULATION OVER 65 AND GOVERNMENT SPENDING ON PENSION BENEFITS*

```{r pressure5, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
library(ggrepel)
pensions <- read_csv('Economist_pensions.csv')
names(pensions) <- c("country","population","spending")
pensions <- pensions %>%
drop_na()

oecd <-subset(pensions,country %in% c("Brazil","Japan","Italy","Turkey","Mexico","China","United States"
                                      ,"Greece", "South Korea", "Poland"))
newentry<- data.frame(country="OECD average", population=mean(oecd$population), spending=mean(oecd$spending))
pensions <- rbind(pensions, newentry)
oecd <- rbind(oecd, newentry)
plot5 <-pensions %>%
  ggplot(aes(x=population, y=spending, label=country))+
  geom_point(colour=ifelse(pensions$country %in% oecd$country, 'red','#ec7063'), size=3)+
  geom_text_repel(data = oecd, fontface=ifelse(oecd$country=="Brazil","bold","plain"))+
  labs(x= "Population aged 65 years and over, % of total", y="Government Spending on pension\nbenefits- % of total GDP", 
       title= "Brazil's Golden Oldie Blowout", subtitle = "Latest available",caption = "Sources: OECD, WorldBank, Previdencia Social")+
  theme(plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.1),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=12, face="bold"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        plot.caption = element_text(colour="#434343", size=8))

plot5
```


### SURFEIT OF SURPLUSES
```{r pressure6, echo=FALSE, fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
balance <- read_csv('Economist_eu-balance_tidy.csv')
balance <- balance %>%
drop_na() 
balance_filter<- balance %>%
filter(measure =="Budget balance, euro million") %>%
select(country,year,value)
Current_filter<- balance %>%
  filter(measure =="Current-account balance, euro million") %>%
  select(country,year,value)
balance_others<-balance_filter %>%
  filter(country %in% c("Austria","Belgium","Italy","Ireland","France", "Portugal"))%>%
  select(year,value)%>%
  group_by(year) %>%
  summarise(value=sum(value))%>%
  mutate(country="Others")
Current_others<-Current_filter %>%
  filter(country %in% c("Austria","Belgium","Italy","Ireland","France", "Portugal"))%>%
  select(year,value)%>%
  group_by(year) %>%
  summarise(value=sum(value))%>%
  mutate(country="Others")
Major_balance<-subset(balance_filter,country %in% c("Germany","Netherlands","Greece","Spain"))
balance_final <- rbind(Major_balance, balance_others)
Major_current<-subset(Current_filter,country %in% c("Germany","Netherlands","Greece","Spain"))
current_final <- rbind(Major_current, Current_others)

plot6a <-balance_final %>%
  ggplot(aes(x=year, y=value, colour=country))+
  geom_bar(stat = 'identity', aes(fill=country))+
  scale_colour_brewer(type = 'qual', palette = 4)+
  labs(x= NULL, y= NULL,title= "Surfeit of Surpluses\nSelected Euro area countries $bn",
       subtitle= "Budget Balances",caption = "Sources: Eurostats, National Statistics")+
  theme(legend.position="bottom", legend.box = "horizontal",
        legend.title = element_blank(),
        legend.background = element_rect(fill = '#ffef99'),
        legend.key.size=unit(0.2, "cm"),
        plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.1),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=11, face="bold"),
        plot.caption = element_text(colour="#434343", size=8))+
         scale_y_continuous(labels = unit_format(scale = 1e-3, unit = ''),
                     breaks = seq(-600000, 10000, 100000),
                     limits = c(-600000, 40000))+
          scale_x_continuous(breaks = seq(2009, 2015, 1))+
          scale_colour_brewer(type = 'qual', palette = 3)

plot6a

plot6b <-current_final %>%
  ggplot(aes(x=year, y=value, colour=country))+
  geom_bar(stat = 'identity', aes(fill=country))+
  scale_colour_brewer(type = 'qual', palette = 4)+
  labs(x= NULL, y= NULL,title= "Surfeit of Surpluses\nSelected Euro area countries $bn",
       subtitle= "Current Account Balances",caption = "Sources: Eurostats, National Statistics")+
  theme(legend.title = element_blank(),
        legend.position="bottom", legend.box = "horizontal",
        legend.background = element_rect(fill = '#ffef99'),
        legend.key.size=unit(0.2, "cm"),
        plot.background = element_rect(fill = '#ffef99'),
        panel.background = element_rect(fill = '#ffd966'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = 'grey25', size=0.1),
        panel.grid.major.y = element_line(colour = 'grey25',size=0.05),
        plot.title = element_text(size=11, face="bold"),
        plot.caption = element_text(colour="#434343", size=8))+
        scale_y_continuous(labels = unit_format(scale = 1e-3, unit = ''),
                     breaks = seq(-150000, 400000, 100000),
                     limits = c(-150000, 400000))+
         scale_x_continuous(breaks = seq(2009, 2015, 1))+
         scale_colour_brewer(type = 'qual', palette = 3)

plot6b
```

###CHART COMMENTS

**Average No. of likes per Facebook post**
•	Improper Scale : Not encouraging the eye to compare different pieces of data
•	Unnecessary variance in bar chart colours
•	Background and white lined grids could be given more subtle colour to emphasize more on the bar graph

**Dog’s measurement**
•	Using Dual Scales on a single plot area is not a good idea
•	Background could be given more subtle colour to emphasize more on the bar graph

**Opinion on Brexit**
•	Using line chart not recommended here as it should be a scatter plot. (the poll data is not continuous)
•	Further, a smoothened line may be drawn to reveal the data at several levels of detail, from a broad overview to the fine structure

**Free Markets & free Workers** 
•	Again, Using Dual Scales on a single plot area is not a good idea
•	Particularly wrong in this case, where y axis of one plot is in negative range, and for the other in positive range

**Population Over 65 and Government Spending on Benefits**
•	Inappropriate variance in country colours. While, to keep graph area clean, only a selected ones have been labelled and contrasted against Brazil, the dual variation in colour tone is a misrepresentation.
•	The ‘latest available’ could be better labelled by the year of data source

**Surfeit of Surpluses**	
•	Very difficult to comprehend due to a number of small categories, many of which are not easy to distinguish because of small value and colour similarities


